Rotation effects on the Lyman-$\alpha$ line morphology in distant galaxies

### 2013-06-10 paper/Analisis.tex
some words on the average number of scatterings

\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}  
\newcommand{\zz}{$z\sim 3$} 
\newcommand{\apj}{ApJ}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}    
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha~}\else{Ly$\alpha$~}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\hs}{{\hspace{1mm}}}  
\newcommand{\kms}{{\ifmmode{{\mathrm{\,km\ s}^{-1}}}\else{\,km~s$^{-1}$}\fi}}

@@

Figure \ref{fig:maximumsvsvelocity} shows the position for the peak
maxima as a function of the rotational velocity $V_{\rm max}$. This
figure shows us clearly that in the case of central distributed
sources there is barely any change with rotational velocity, at least
in the range of explirea parameters. However, in the case of
homogeneously emitted sources the maxima position remain close to
constant until beyond some velocity threshold the line becomes single
peaked with $x_{\rm max}=0$.

The transition to a single peak line most probably occurs for systems
where it  becomes easier for the bulk of the photons to escape with the lowest
number of scatterings possible, allowing them not to move very far
from the center of the line. This hypothesis can also help to explain how the
single peak stage is easily achieved in the homogeneous source
distribution compared to the central distribution. In the homogeneous
case there is a fraction of the photons that are inside a photosphere region with
$\tau_{H,r}\ll \tau_{H}$ where $\tau_{H,r}$ is the optical depth from
the radius of emission to the sphere's surface. This conditions allows
the photons to escape with much less scatterings compared to the
photons emitted at the very center of the sphere. In turn, it gives
the photons less scatterings to be placed far from the line
center. Increasing the rotational velocity $V_{\rm max}$ reduces the
optical depth making the  photosphere region effectively larger,
increasing the number of photons escaping close to the lines's center.

For larger values of $\tau_{H}$ the transitional velocity $V_{\rm
  trans}$ is also larger. In our models we find the following
correspondance between the optical depth $\tau_{\rm H}=\{10^5, 10^6,
10^7\}$ and the transitional velocities $V_{\rm
  trans}=\{50-100\kms, 100-200\kms, 200-300\kms\}$ which can only be
constrained to be in the range of velocities in the models that gave a $x_{\rm
  max}\neq 0$ and $x_{\rm max}=0$.

@@

Until know we have focus in how rotation affect the morphology of the 
Lyman alpha line, now we turn to study deeply the possible causes of 
this effects. As recombination is the main cause of the emission of the 
Lyman alpha line, it is important to study the relationship between the 
number of times that a photon is absorbed and re emitted that we simply 
take as scatterings with it's shift in the wavelength.

As a first approach we study the number of scatterings of all the photons 
at different velocities and for both photon distributions Figure~\ref{fig:Nscatt}. 
For the central distribution we found that the average number of scattering 
does not have a relevant change ($\sim 0.5\%$) with $V_{max}$, For the 
Homogeneous distribution we found that as $V_{max}$ increase the average 
number of scatterings decrease in a$\sim 39\%$, it means that rotation 
makes photons to escape easier from the cloud. The reason why this is not 
observed in the central distribution is because the amount of gas that every 
photon has to break through is so large that the average number of scatterings 
is large too, making the effect of rotation unobservable.

>>>>

\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}  
\newcommand{\zz}{$z\sim 3$} 
\newcommand{\apj}{ApJ}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}    
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\nscatt}{\langle N_{\rm  scatt}\rangle}
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha~}\else{Ly$\alpha$~}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\hs}{{\hspace{1mm}}}  
\newcommand{\kms}{{\ifmmode{{\mathrm{\,km\ s}^{-1}}}\else{\,km~s$^{-1}$}\fi}}

@@

Figure \ref{fig:maximumsvsvelocity} shows the position for the peak
maxima as a function of the rotational velocity $V_{\rm max}$. This
figure shows us clearly that in the case of central distributed
sources there is barely any change with rotational velocity in the
range of explored parameters. However, in the case of
homogeneously emitted sources the maxima position remain close to
constant until beyond some velocity threshold the line becomes single
peaked with $x_{\rm max}=0$.

It is probable that the transition to a single peak line occurs for systems
where it  becomes easier for the bulk of the photons to escape with the lowest
number of scatterings possible, allowing them not to move very far
from the center of the line. This hypothesis can help to explain how the
single peak stage is easily achieved in the homogeneous source
distribution. In this case there is a fraction of the photons that are
inside a photosphere region with $\tau_{H,r}\ll \tau_{H}$ where
$\tau_{H,r}$ is the optical depth from the radius of emission to the
sphere's surface. This conditions allows the photons to escape with
much less scatterings compared to the photons emitted at the very
center of the sphere. In turn, it gives the photons less scatterings
to be placed far from the line center. Increasing the rotational
velocity $V_{\rm max}$ reduces the optical depth making the
photosphere region effectively larger, increasing the number of
photons escaping close to the lines's center. For the central case the
photosphere is not present, and other mechanisms must be at play in
the steady reduction of the double peak.

For larger values of $\tau_{H}$ the transitional velocity $V_{\rm
  trans}$ is also larger. In our models we find the following
correspondance between the optical depth $\tau_{\rm H}=\{10^5, 10^6,
10^7\}$ and the transitional velocities $V_{\rm
  trans}=\{50-100\kms, 100-200\kms, 200-300\kms\}$ which can only be
constrained to be in the range of velocities in the models that gave a $x_{\rm
  max}\neq 0$ and $x_{\rm max}=0$.   In the next section explore the
origin of this trends and consider to what extent the hypothesis of a low
number of scatterings is correlated with the emergence of a single
peak.

@@

The number of times that a \ly photon is absorbed and re-emmited is
connected to the final frequency that it can have after escaping the
galaxy. In the case of static gas geometries, a large value of the
optical depth is immediatly followed by a high number of
scatterings. In turn a large optical depth increases the probability
that a \lya photon will be found far from the center of the line. The
peak maxima shifting away form the line center as the amount of
neutral hydrogen increases.

In Figure \ref{fig:Nscatt} we see the average number of scatterings
$\langle N_{\rm scatt}\rangle$ as a function of the rotational velocity
$V_{\rm max}$. For the central distributions we find that there is not
a significante change for increasing rotational
velocities, $\langle N_{\rm scatt}\rangle$ changes less than $0.5\%$
for different velocities. In this case we also find that the average
number of scatterings is proportional to the optical depth, as
expected in analogy from the analytic result for the homogeneous
infinite-slab $\langle N_{\rm  scatt}\rangle=1.612\tau_{\rm   H}$
\citep{Adams72,Harrington73}. In our experiments we find  $\langle N_{\rm
  scatt}\rangle= XXX\tau_{\rm   H}$.

However, for the homogeneous distribution there is a clear decrease of
$\langle N_{\rm  scatt}\rangle$ as the  $V_{\rm max}$ increases. This
effect more pronounced for the lower values of the optical depth. For
$\tau_{\rm H}=10^5$ the average number of scatterings decreases by
$XX\%$ at $V_{\rm max}=300\kms$ in comparison to the static case.

The analytic expectation for the slab with homogeneously emitted
sources is $\langle N_{\rm  scatt}\rangle=1.16\tau_{\rm   H}$
\citep{Harrington73}, a factor of $0.72$ lower than the case of the
centrally emitted photons. In our case we find that for the static
spheres this factor is $XXX$.

%%%%%
